---
title: "Untitled"
output: html_document
date: "`r Sys.Date()`"
---

```{r Libraries}
library(ggplot2)
```

### Progress so far: fully understand the given dataset, made 1 spike train raster plot of neuron 1 in trial 1 in session 1 for the mouse Cori, created structures and details to put in each section of the report.

# Abstract

# Introduction

# Background

In the study conducted by Steinmetz et al. (2019), experiments were performed on a total of 10 mice over 39 sessions, each composed of several hundred trials. Among the 39 sessions, 3 sessions (526 trials in total) were from the mouse named Cori and 4 sessions (904 trials in total) were from Forssmann. The analysis done in this project focuses specifically on the data from the 3 sessions on Cori and 2 sessions on Forssmann. From supplementary Table 2 of Steinmetz et al. (2019), all 3 sessions for Cori contained VIS-related regions. Only session from 2017-11-01 and 2017-11-02 contain recordings from VIS for Forssmann. TBU

# Data Structure 

In session 1 of Cori, 214 trials of experiment were performed. In each trial, information of the firing conditions of 178 neurons were collected. For each neuron, 39 timestamps were recorded. At each timestamp, the corresponding fire/no-fire status of the neuron was also recorded. A fire corresponds to the number 1 and vice versa in record. With the information, we can generate a spike raster plot (Fig.1) for neuron 1 in trial 1 from time 65.27941 second (the onset of the stimulus) to time 65.65941 second (0.38s or roughly 0.4s post-onset).

## Requirements:
Review & provide basic background of the experiment. (More on practice project outline)

Experiment description:
10 mice
39 sessions (?QUESTIONS? WHAT ARE THE SESSIONS HERE?)
    Each session comprised several hundred trials, same 1 mice
        Each trial, visual stimuli with contrast levels, 1 left & 1 right screen, random
            Stimuli contrast levels on each screen: 0 (absense), 0.25, 0.5, 1
            Mice, forepaw controlled wheel, decide -> outcome -> reward/penalty
            Record neUron activity in visual cortex -> spike trains (?QUESTION? WHATS THIS?) -> collections of timestamps == neuron firing
            
Project Scope:
Spike trains from onset of the stimuli to 0.4s post-onset
Session 1-5, mice Cori and Frossman

Sample size:
n=10 mice were used because this was a sufficient number to record n=29134 neurons in n=42 brain regions, numbers which are greater than any previously published work. No statistical calculations were used to determine exactly how many neurons to record, i.e. the counts are exact.

Data exclusion: (?QUESTION? WHAT DOES THIS MEAN?)
Behavioral trials were excluded when there were three or more miss trials in a row indicating disengagement; this criterion was pre- established. Neurons were excluded when the null hypothesis that their firing rate was un-modulated during the task could not be rejected according to a series of four tests for different types of modulation; this criterion was not determined in advance but was developed from data inspection.

Replication:
Recordings were undertaken in n=10 mice and a population of neurons recorded in each brain region in at least two of the mice.

Randomization:
Trial types during the behavioral task were interleaved randomly.

7 variables:
1. mouse_name
2. date_exp
3. feedback_type: 1 (succ) & -1 (fail) (?QUESTION? HOW TO DEFINE THIS FEEDBACK?)
4. contrast_left: left stimulus contrast
5. contrast_right: 
6. time: centers of the time bins for spks (?QUESTION? WHAT DOES THIS MEAN? WHY DUPLICATE TIMES?)
7. spks: num spikes in time bins defined in `time` (?QUESTION? HOW TO INTERPRET THIS VARIABLE? WHAT IS THE ROW AND COUMN NUM? ALL SESS HAVE 39 COL BUT DIFF ROW NUM.)

Questions of Interest:
1. How neurons respond to left and right stml? 
    ?QUESTION? 
    HOW TO DEFINE THIS HOW?
    DID IT RESPOND? 
    HOW EARLY THE FIRST NEURON? 
    HOW MANY NEURONS? 
    HOW FAST THE WHOLE TRANSMISSION PROCESS BTW NEURONS?
    DIFF RESP FOR LEFT AND RIHGT? OR SAME?
2. How to predict outcome of each trial using neural activities & stimuli?

Method for Q1:
- Formulate statistically sound hypotheses. 
- Does the l n r stml have additive effects on the neural responses? 
    i.e. whether the interaction effect exists.
    model & hypo in outline

Method for Q2:
- Only goal of pred md: the best pred performance. Eval by the SENSITIVITY and SPECIFICITY evaluated on the first 100 trials in session 1.
- Enhance the pred performance by throughly analyzing the reward mechanism described in Steinmetz et al. (2019).

# Descriptive Analysis

?QUESTION? DOES DATE OF THE EXPERIMENT MATTER? 

## Requirements:
1. Summary stats, explain findings
2. Plots, explain findings
3. !IMPORTANT! Address the unique feature of this dataset, which is that each session contains varying numbers of neurons (?QUESTION? HOW TO SEE VARYING NUM OF NEURONS?)
4. !IMPORTANT! Define the outcome var
    Suggested approach: use the mean firing rate = for each trial, the avg num spikes per sec across all neurons within the 0.4s interval.
    !IMPORTANT! Justify this choice of mean firing rate.
    
## Data Structure    
39 sessions in total for 10 mice, 3 sessions specifically for mouse Cori
Session 1 among the 3 session for Cori contains 214 trials, for each trial:
    39 timestamps
    there are 178 neurons
    
Spike raster plot:
    trial_id vs time for each neuron
    nueron_id vs time for each trial
    
```{r Provided Sample Code from Course Project Description Rmd file}
#TODO# Add in reference section if this code is used

session = list()
for (i in 1:5) {
  session[[i]] = readRDS(paste('data/session', i, '.rds', sep = ''))
  print(session[[i]]$mouse_name)
  print(session[[i]]$date_exp)
}

# Take the 11th trial in Session 1 for example
id=11
session[[1]]$feedback_type[id]
session[[1]]$contrast_left[id]
session[[1]]$contrast_right[id]
length(session[[1]]$time[[id]])
dim(session[[1]]$spks[[id]])

# Obtain the firing rate 
# averaged over [0,0.4] seconds since stim onsets
# averaged across all neurons 

ID=1
t=0.4 # from Background 

n.trials=length(session[[ID]]$spks)
n.neurons=dim(session[[ID]]$spks[[1]])[1]

# Obtain the firing rate 
firingrate=numeric(n.trials)
for(i in 1:n.trials){
  firingrate[i]=sum(session[[ID]]$spks[[i]])/n.neurons/t
}
```


```{r Spike trian raster plot}
# Session 1 for the mouse Cori
ID = 1
t = 0.4 # from Background 

n_trials = length(session[[ID]]$spks)
n_neurons = dim(session[[ID]]$spks[[1]])[1]

range_trials = 1:10
range_neurons = 1:5

session_1 = session[[ID]]


trial_1_time = session_1$time[[1]]
trial_1_time_from_stimulus = trial_1_time - trial_1_time[1]
trial_1_neuron_1_spikes = session_1$spks[[1]][1,]

temp_neuron_1_data = data.frame(trial = rep(1, 39),
                                time_from_stimulus = trial_1_time_from_stimulus,
                                spikes = trial_1_neuron_1_spikes)

ggplot(data=temp_neuron_1_data,
       mapping=aes(time_from_stimulus, spikes)) +
    geom_linerange(mapping=aes(ymin=0, 
                               ymax=spikes)) +
    labs(caption="Fig.1 Spike raster plot for neuron 1 in trial 1 for Cori")
```



    
# Inferential Analysis
Address Q1

?QUESTION? SPLIT PLOT WHOLE PLOT DESIGN (PIAZZA)

Factors: left contrast, right contrast, mourse name
outcome: firing rate across all neurons within the given 0.4 (0.38 caclulated from the timestamp difference from the data) seconds time interval.

## Requirements
1. Mixed effect model
- Random effect on the mice, Cori vs Forssmann


            

# Reference

Steinmetz, N.A., Zatka-Haas, P., Carandini, M. et al. Distributed coding of choice, action and engagement across the mouse brain. Nature 576, 266--273 (2019). <https://doi.org/10.1038/s41586-019-1787-x>

Paper Source Code: https://github.com/nsteinme/steinmetz-et-al-2019





# Resources that could be used as reference

Code for plotting spike train raster plots in R:
1. https://gist.github.com/Vindaar/9c32c0676ffddec9078e4c0917861fcd
2. https://rdrr.io/rforge/STAR/man/raster.html
3. https://rdrr.io/github/dpwynne/mmnst/man/RasterPlot.html
4. https://www.imsbio.co.jp/RGM/R_rdfile?f=STAR/man/raster.Rd&d=R_CC

Neural Encoding: http://www.columbia.edu/cu/appliedneuroshp/Spring2018/Spring18SHPAppliedNeuroLec3.pdf

Spike raster plot & Normalized PSTH plot (Python): https://neuraldatascience.io/6-single_unit/intro_spike_trains.html

